Recursive algorithms, branching coefficients and applications

### 2010-09-29 AffineLieAlgebras/InformalNotes/article.tex
Internal references are fixed
Minor changes

\documentclass[12pt]{iopart}
%\usepackage{ucs}
%\usepackage{amsmath,amssymb,amsthm}
\usepackage{iopams,amssymb,amsthm}
%\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage{color}
%\usepackage[pdftex]{graphicx}
\usepackage{graphicx}
%\pagestyle{plain}
%\usepackage[unicode,verbose,plainpages=false]{hyperref}
\usepackage{hyperref}

@@

There exist several approaches to deal with the branching coefficients. Some of them use the BGG
resolution \cite{bernstein1975differential} (for Kac-Moody algebras the algorithm is described in
\cite{kac1990idl},\cite{wakimoto2001idl}), the Schure function series \cite{fauser2006new}, the BRST
cohomology \cite{Hwang:1994yr}, Kac-Peterson formulas \cite{kac1990idl,quella2002branching} or the
combinatorial methods applied in \cite{feigin707principal}.

@@

Consider affine Lie algebras $\frak{g}$ and $\frak{a}$ with the
underlying finite-dimensional subalgebras $\go$ and $%
\ao$ and an injection $\frak{a}\longrightarrow \frak{g%
}$ such that $\frak{a}$ is a reductive subalgebra $\frak{a\subset g}$ with
correlated root spaces: $\frak{h}_{\frak{a}}^{\ast }\subset \frak{h}_{\frak{g%
}}^{\ast }$ and $\frak{h}_{\ao}^{\ast }\subset \frak{h%
}_{\go}^{\ast }$\ .

We use the following notations:

@@

\subsection{Proof of the recurrent relation}
\label{sec:proof}
\begin{theorem}
  For the anomalous branching coefficients (\ref{eq:21}) holds the recurrent relation
  \begin{equation}
    \label{eq:1}
    \begin{array}{c}
      k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
        \sum_{\omega\in W_{\bot}\backslash W} \epsilon(\omega)\;
        {\rm dim}\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right)
        \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+ \right.\\
      \left.
        +\sum_{\gamma \in
          \Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
          +\gamma }^{\left( \mu \right) }\right).
    \end{array}
  \end{equation}
  \begin{proof}
    Consider the branching of a module $L_{\frak{g}}^{\mu }$ in terms of formal characters and
    projection operators $\pi_{\mathfrak{a}}:P\to P_{\mathfrak{a}}$:
    \begin{equation*}
      % \label{eq:3}
      L_{\frak{g}\downarrow \frak{a}}^{\mu }=\bigoplus\limits_{\nu \in P_{\frak{a}%
        }^{+}}b_{\nu }^{\left( \mu \right) }L_{\frak{a}}^{\nu }\quad
      \Longrightarrow\quad
      \pi_{\mathfrak{a}}(ch L^{\mu}_{\mathfrak{g}})=
      \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu} ch L^{\nu}_{\mathfrak{a}}.
    \end{equation*}
    The Weyl-Kac character formula leads to
    the relation
    \begin{equation}
      \label{eq:4}
      \pi_{\mathfrak{a}}\left(\frac{\sum_{\omega\in W} \epsilon(\omega) e^{\omega(\mu+\rho)-\rho}}
        {\prod_{\alpha\in\Delta^{+}}(1-e^{-\alpha})^{\mathrm{mult}(\alpha)}}\right) =
      \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu}
      \frac{\sum_{\omega\in W_{\mathfrak{a}}}\epsilon(\omega)
        e^{\omega(\nu+\rho_{\mathfrak{a}})-\rho_{\mathfrak{a}}}}
      {\prod_{\beta\in \Delta_{\mathfrak{a}}^{+}}(1-e^{-\beta})^{\mathrm{mult}_{\mathfrak{a}}(\beta)}}.
    \end{equation}

@@

    Using the definition \ref{fan-definition} we finally get the recurrent relation
    \begin{eqnarray*}
      k_{\xi }^{\left( \mu \right) }=
      -\frac{1}{s\left( \gamma _{0}\right) }
      \left(
        \sum_{\omega\in W_{\bot}\backslash W} \epsilon(\omega)\; \mathrm{dim}
        \left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right)
        \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+
      \right.\\
      \left. +
        \sum_{\gamma \in \Gamma _{\frak{a}\subset \frak{g}}} s\left( \gamma +\gamma _{0}\right) k_{\xi+\gamma }^{\left( \mu \right) }
      \right),
      \label{recurrent-relation}
    \end{eqnarray*}
\end{proof}
\end{theorem}
\begin{corollary}
  Since $b^{(\mu)}_{\nu}=k^{(\mu)}_{\nu}\;\mbox{for}\;\nu\in \bar{C_{\mathfrak{a}}}$ the recurrent relation gives the values of the branching coefficients inside the main Weyl chamber.
\end{corollary}
\begin{corollary}
  Consider the case  $\Delta_{\bot}^{+} = 0$. There are three different reasons for $\Delta_{\bot}^{+}$
to be empty: i) $\mathrm{dim}\mathfrak{h}_{\mathfrak{a}}=\mathrm{dim}\mathfrak{h}_{\mathfrak{g}}$,
ii) $\mathfrak{a}_{\bot}=0$ and iii) $\mathfrak{a}_{\bot}\subset \mathfrak{h}_{\mathfrak{g}}$.
Both the first and the second cases can be treated as corresponding to the trivial orthogonal subalgebra:
$\mathfrak{a}_{\bot}=0$. In any of these cases the formal characters in the right-hand side of (\ref{eq:13})
are reduced to the element $e^{ \pi_{\mathfrak{a}_{\bot}}\omega(\mu+\rho)}$.
In the first two cases the vector $\omega(\mu+\rho)$ is projected to the subspace orthogonal to the weight space of $\mathfrak{a}$. It is clear that in any of the three variants the final vector $\pi_{\mathfrak{a}}\pi_{\mathfrak{a}_{\bot}}\omega(\mu+\rho)$ leads to the unit of the formal algebra
$\mathcal{E}$. Thus when the set $\Delta_{\bot}^{+}$ is empty the  recurrent relation is simplified:
\begin{equation*}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{w\in W}\epsilon \left( w\right) \delta _{\xi ,\pi _{\frak{a}}\circ
\left( w\circ (\mu +\rho )-\rho \right) +\gamma _{0}}+\sum_{\gamma \in
\Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
+\gamma }^{\left( \mu \right) }\right),  \label{recurrent relation special}
\end{equation*}
the latter coincides with the one obtained in \cite{ilyin812pbc} (formula (16)).
\end{corollary}
\subsection{Algorithm for recursive computation of branching coefficients}
\label{sec:algorithm}

@@

The substitution of the generators of $\mathfrak{a}$  in terms of the generators of $\mathfrak{g}$ into these combinations  should give the energy-momentum tensor $T_{\mathfrak{g}}$. But if the set of the generators $\Delta_{\bot}$ is not empty this is not possible, since $T_{\mathfrak{g}}$ contains the combinations of the generators $X^{\alpha}_n, \; \alpha\in \Delta_{\bot}$.
\end{proof}
\end{mynote}
\subsubsection{Special embedding $\hat{A}_1\subset\hat{A}_2$}
\label{sec:spec-embedd-hata_1s}
Consider the case where both $\mathfrak{g}$ and $\mathfrak{a}$ are affine Lie algebras:
$\hat{A}_1 \longrightarrow \hat{A}_2$ and the injection is the affine extension of the
special injection $A_1 \longrightarrow A_2$ with the embedding index $x_e=4$.
As far as the $\mathfrak{g}$-modules to be considered are of level one,
the $\mathfrak{g}$-modules will be of level $\tilde{k}=kx_e=4$.

>>>>

\documentclass[12pt]{iopart}
%\usepackage{ucs}
%\usepackage{amsmath,amssymb,amsthm}
\usepackage{iopams,amssymb,amsthm}
%\usepackage{pb-diagram}
\usepackage{multicol}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian]{babel}
\usepackage{cmap}
\usepackage{color}
\usepackage[pdftex]{graphicx}
\usepackage{epstopdf}
%\usepackage{graphicx}
%\pagestyle{plain}
%\usepackage[unicode,verbose,plainpages=false]{hyperref}
\usepackage{hyperref}

@@

There exist several different approaches to deal with the branching coefficients. Some of them use the BGG
resolution \cite{bernstein1975differential} (for Kac-Moody algebras the algorithm is described in
\cite{kac1990idl},\cite{wakimoto2001idl}), the Schur function series \cite{fauser2006new}, the BRST
cohomology \cite{Hwang:1994yr}, Kac-Peterson formulas \cite{kac1990idl,quella2002branching} or the
combinatorial methods applied in \cite{feigin707principal}.

@@

Consider affine Lie algebras $\frak{g}$ and $\frak{a}$ with the
underlying finite-dimensional subalgebras $\go$ and $%
\ao$ and an injection $\frak{a}\longrightarrow \frak{g%
}$ such that $\frak{a}$ is a reductive subalgebra $\frak{a\subset g}$ with
correlated root spaces: $\frak{h}_{\frak{a}}^{\ast }\subset \frak{h}_{\frak{g%
}}^{\ast }$ and $\frak{h}_{\ao}^{\ast }\subset \frak{h%
}_{\go}^{\ast }$\ 
.
We use the following notations:

@@

\subsection{Proof of the recurrent relation}
\label{sec:proof}
\begin{theorem}
  For the anomalous branching coefficients (\ref{eq:21}) holds the recurrent relation
  \begin{equation}
    \label{eq:1}
    \label{recurrent-relation}
    \begin{array}{c}
      k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
        \sum_{\omega\in W_{\bot}\backslash W} \epsilon(\omega)\;
        {\rm dim}\left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right)
        \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+ \right.\\
      \left.
        +\sum_{\gamma \in
          \Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
          +\gamma }^{\left( \mu \right) }\right).
    \end{array}
  \end{equation}
  \begin{proof}
    Consider the branching of a module $L_{\frak{g}}^{\mu }$ in terms of formal characters and
    projection operators $\pi_{\mathfrak{a}}:P\to P_{\mathfrak{a}}$:
    \begin{equation*}
      % \label{eq:3}
      L_{\frak{g}\downarrow \frak{a}}^{\mu }=\bigoplus\limits_{\nu \in P_{\frak{a}%
        }^{+}}b_{\nu }^{\left( \mu \right) }L_{\frak{a}}^{\nu }\quad
      \Longrightarrow\quad
      \pi_{\mathfrak{a}}(ch L^{\mu}_{\mathfrak{g}})=
      \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu} ch L^{\nu}_{\mathfrak{a}}.
    \end{equation*}
    The Weyl-Kac character formula leads to
    the relation
    \begin{equation}
      \label{eq:4}
      \pi_{\mathfrak{a}}\left(\frac{\sum_{\omega\in W} \epsilon(\omega) e^{\omega(\mu+\rho)-\rho}}
        {\prod_{\alpha\in\Delta^{+}}(1-e^{-\alpha})^{\mathrm{mult}(\alpha)}}\right) =
      \sum_{\nu\in P^{+}_{\mathfrak{a}}}b^{(\mu)}_{\nu}
      \frac{\sum_{\omega\in W_{\mathfrak{a}}}\epsilon(\omega)
        e^{\omega(\nu+\rho_{\mathfrak{a}})-\rho_{\mathfrak{a}}}}
      {\prod_{\beta\in \Delta_{\mathfrak{a}}^{+}}(1-e^{-\beta})^{\mathrm{mult}_{\mathfrak{a}}(\beta)}}.
    \end{equation}

@@

    Using the definition \ref{fan-definition} we finally get the recurrent relation
    \begin{eqnarray*}
      k_{\xi }^{\left( \mu \right) }=
      -\frac{1}{s\left( \gamma _{0}\right) }
      \left(
        \sum_{\omega\in W_{\bot}\backslash W} \epsilon(\omega)\; \mathrm{dim}
        \left(L^{\pi_{\mathfrak{a}_{\bot}}(\omega(\mu+\rho))-\rho_{\mathfrak{a}_{\bot}}}_{\mathfrak{a}_{\bot}}\right)
        \delta_{\xi-\gamma_0,\pi_{\mathfrak{a}}(\omega(\mu+\rho)-\rho)}+
      \right.\\
      \left. +
        \sum_{\gamma \in \Gamma _{\frak{a}\subset \frak{g}}} s\left( \gamma +\gamma _{0}\right) k_{\xi+\gamma }^{\left( \mu \right) }
      \right),
    \end{eqnarray*}
\end{proof}
\end{theorem}
\begin{corollary}
  Since $b^{(\mu)}_{\nu}=k^{(\mu)}_{\nu}\;\mbox{for}\;\nu\in \bar{C_{\mathfrak{a}}}$ the recurrent relation gives the values of the branching coefficients inside the main Weyl chamber.
\end{corollary}
\begin{corollary}
  Consider the case  $\Delta_{\bot}^{+} = 0$. There are three different reasons for $\Delta_{\bot}^{+}$
to be empty: i) $\mathrm{dim}\mathfrak{h}_{\mathfrak{a}}=\mathrm{dim}\mathfrak{h}_{\mathfrak{g}}$,
ii) $\mathfrak{a}_{\bot}=0$ and iii) $\mathfrak{a}_{\bot}\subset \mathfrak{h}_{\mathfrak{g}}$.
Both the first and the second cases can be treated as corresponding to the trivial orthogonal subalgebra:
$\mathfrak{a}_{\bot}=0$. In any of these cases the formal characters in the right-hand side of (\ref{eq:13})
are reduced to the element $e^{ \pi_{\mathfrak{a}_{\bot}}\omega(\mu+\rho)}$.
In the first two cases the vector $\omega(\mu+\rho)$ is projected to the subspace orthogonal to the weight space of $\mathfrak{a}$. It is clear that in any of the three variants the final vector $\pi_{\mathfrak{a}}\pi_{\mathfrak{a}_{\bot}}\omega(\mu+\rho)$ leads to the unit of the formal algebra
$\mathcal{E}$. Thus when the set $\Delta_{\bot}^{+}$ is empty the  recurrent relation is simplified:
\begin{equation*}
k_{\xi }^{\left( \mu \right) }=-\frac{1}{s\left( \gamma _{0}\right) }\left(
\sum_{w\in W}\epsilon \left( w\right) \delta _{\xi ,\pi _{\frak{a}}\circ
\left( w\circ (\mu +\rho )-\rho \right) +\gamma _{0}}+\sum_{\gamma \in
\Gamma _{\frak{a}\subset \frak{g}}}s\left( \gamma +\gamma _{0}\right) k_{\xi
+\gamma }^{\left( \mu \right) }\right),  \label{recurrent relation special}
\end{equation*}
the latter coincides with the one obtained in \cite{ilyin812pbc} (formula (16)).
\end{corollary}
\subsection{Algorithm for recursive computation of branching coefficients}
\label{sec:algorithm}

@@

The substitution of the generators of $\mathfrak{a}$  in terms of the generators of $\mathfrak{g}$ into these combinations  should give the energy-momentum tensor $T_{\mathfrak{g}}$. But if the set of the generators $\Delta_{\bot}$ is not empty this is not possible, since $T_{\mathfrak{g}}$ contains the combinations of the generators $X^{\alpha}_n, \; \alpha\in \Delta_{\bot}$.
\end{proof}
\end{mynote}
\subsubsection{Special embedding $\hat{A}_1\subset\hat{A}_2$}
\label{sec:spec-embedd-hata_1s}

Consider the case where both $\mathfrak{g}$ and $\mathfrak{a}$ are affine Lie algebras:
$\hat{A}_1 \longrightarrow \hat{A}_2$ and the injection is the affine extension of the
special injection $A_1 \longrightarrow A_2$ with the embedding index $x_e=4$.
As far as the $\mathfrak{g}$-modules to be considered are of level one,
the $\mathfrak{g}$-modules will be of level $\tilde{k}=kx_e=4$.
